---
title: "Military Operations"
subtitle: "A webpage for the Military Operations with Novel Strategic Technologies (MONSTr) Dataset"
page-layout: full
format:
  html:
    toc: true
    toc-location: right
    html-math-method: katex
    css: styles.css
    citations-hover: true
    footnotes-hover: true
    code-tools: true
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: true
editor: visual
execute:
  echo: false
  warning: false
  cache: true
---


```{r knitr_options, echo = FALSE, warning = FALSE, cache = FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
```

```{r}
# Prep data
## Load main data
df <- read.csv(paste0(here::here(), "/content/data/","MONSTr.csv"))
```

## Fine-grained data on US military operations

Scraping wikidata and DBpedia, we identify a number of covariates about US military operations since 1989, including their location, participants, start and end date, and military means the US used.

```{r}
# Prep actors for the map
## Load raw dbpedia data that includes all actors
df_dbpedia <- read.csv(paste0(here::here(), "/content/data/","02a_interventions_newdata_dbpedia.csv")) %>%
  dplyr::select(wikidata_id, belligerent_sideA, belligerent_sideB) %>%
  dplyr::rename(belligerent1 = belligerent_sideA,
                belligerent2 = belligerent_sideB) %>%
  dplyr::distinct()

df_sideA <- df_dbpedia %>%
  dplyr::select(wikidata_id, belligerent1) %>%
  dplyr::distinct() %>%
  dplyr::group_by(wikidata_id) %>%
  dplyr::mutate(sideA = paste(belligerent1, collapse = ", ")) %>%
  dplyr::ungroup() %>%
  dplyr::select(wikidata_id, sideA) %>%
  dplyr::distinct()

df_sideB <- df_dbpedia %>%
  dplyr::select(wikidata_id, belligerent2) %>%
  dplyr::distinct() %>%
  dplyr::group_by(wikidata_id) %>%
  dplyr::mutate(sideB = paste(belligerent2, collapse = ", ")) %>%
  dplyr::ungroup() %>%
  dplyr::select(wikidata_id, sideB) %>%
  dplyr::distinct()

df <- dplyr::left_join(df, df_sideA)
df <- dplyr::left_join(df, df_sideB)

df_sf <- df %>% dplyr::filter(!is.na(lat)) %>%
  dplyr::filter(!is.na(lon))

## New column for means used
df_sf$means_aerialbombinglab[df_sf$means_aerialbombing == 1] <- "aerial bombing"
df_sf$means_airtoairlab[df_sf$means_airtoair == 1] <- "air to air"
df_sf$means_closeairsupportlab[df_sf$means_closeairsupport == 1] <- "close air support"
df_sf$means_cruisemissileslab[df_sf$means_cruisemissiles == 1] <- "cruise missiles"
df_sf$means_droneslab[df_sf$means_drones == 1] <- "drone strikes"
df_sf$means_groundtroopslab[df_sf$means_groundtroops == 1] <- "ground troops"
df_sf$means_paramilitarylab[df_sf$means_paramilitary == 1] <- "paramilitary"

df_sf$meanslab <- paste0(df_sf$means_aerialbombinglab, ", ", 
                         df_sf$means_airtoairlab, ", ", 
                         df_sf$means_closeairsupportlab, ", ", 
                         df_sf$means_cruisemissileslab, ", ", 
                         df_sf$means_droneslab, ", ", 
                         df_sf$means_groundtroopslab, ", ", 
                         df_sf$means_paramilitarylab)

# Delete NA and ", ," and padding
df_sf$meanslab <- gsub("NA, ", "", df_sf$meanslab)
df_sf$meanslab <- gsub(", NA", "", df_sf$meanslab)

# Create pop up label
## Pop up formatting
sep <- "<br>"
close_sep <- "</br>"
str_open <- "<strong>"
str_close <- "</strong>"

## Create label column
df_sf$label <- paste0(sep, str_open, "Operation: ", str_close, df_sf$wikidata_name, close_sep, 
                      sep, str_open, "Part of: ", str_close, df_sf$parent_name, close_sep, 
                      sep, str_open, "Dates: ", str_close, df_sf$start_date, " to ", df_sf$end_date, close_sep,
                      sep, str_open, "Side A: ", str_close, df_sf$sideA, close_sep,
                      sep, str_open, "Side B: ", str_close, df_sf$sideB, close_sep,
                      sep, str_open, "Means: ", str_close, df_sf$meanslab, close_sep)

# Coerce to sf object
df_sf <- sf::st_as_sf(df_sf, coords = c("lon", "lat"))

# Calculate center for view
center_lon <- mean(df$lon, na.rm = TRUE)
center_lat <- mean(df$lat, na.rm = TRUE)

# Plot with leaflet
leaflet::leaflet() %>%
  leaflet::addTiles() %>%
  leaflet::addCircleMarkers(data = df_sf, popup = df_sf$label, clusterOptions = leaflet::markerClusterOptions()) %>%
  leaflet::setView(center_lon, center_lat, zoom = 2)
```

## Structure relationships

We also identify the nested relationships among US operations. Which operations are a part of which broader operations? Under what wars do each of these operations fall? We match these data to existing interventions datasets that contain information about the broader interventions in which these operations occurred.

```{r}
read.csv(paste0(here::here(),"/content/data/tree_ops.csv")) %>%
  dplyr::filter(stringr::str_detect(row_operation, "_Q")) %>%
  dplyr::mutate(dplyr::across(everything(), stringr::str_remove_all, pattern = '_.*')) %>%
  collapsibleTree::collapsibleTree(hierarchy = c("level_2", "level_3", "level_4", "level_5", "level_6", "level_7", "level_8", "level_9"), 
                                   height = 20000, width = 8000, 
                                   collapse = FALSE, zoomable = FALSE)
```
