---
title: "Military Operations with Novel Strategic Technologies (MONSTr) Dataset"
page-layout: full
format:
  html:
    html-math-method: katex
    css: styles.css
    citations-hover: true
    footnotes-hover: true
editor: visual
execute:
  echo: false
  warning: false
  cache: false
---

```{r knitr_options}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
```

```{r}
# Prep data
## Load main data
df <- read.csv(paste0(here::here(), "/content/data/","MONSTr.csv"))
```

## Purpose

This website introduces a comprehensive dataset on the universe of United States military operations from 1989-2021 from a single source: Wikipedia. Using automated extraction techniques on its two structured knowledge databases$-$Wikidata and DBpedia$-$we uncover information about individual operations within nearly every post-1989 military intervention described in existing academic datasets. The data we introduce offers unprecedented coverage and granularity that enables analysis of myriad factors associated with when, where, and how the United States employs military force.

MONSTr covers operation-level events occurring within the auspices of a military intervention. Each observation thus represents:

*A series of tactical actions (battles, engagements, strikes) conducted by combat forces in an operational theater to achieve strategic or campaign objectives in the context of a political issue or dispute through action against a foreign adversary. Routine military movements and operations without a defined target like military training exercises, the routine forward deployment of military troops, non-combatant evacuation operations, and disaster relief are excluded.*

The interactive map below shows the location of each US military operation. Zoom in or click on a node to get more information about a specific event.

```{r}
# Prep actors for the map
## Load raw dbpedia data that includes all actors
df_dbpedia <- read.csv(paste0(here::here(), "/content/data/","02a_interventions_newdata_dbpedia.csv")) %>%
  dplyr::select(wikidata_id, belligerent_sideA, belligerent_sideB) %>%
  dplyr::rename(belligerent1 = belligerent_sideA,
                belligerent2 = belligerent_sideB) %>%
  dplyr::distinct()

df_sideA <- df_dbpedia %>%
  dplyr::select(wikidata_id, belligerent1) %>%
  dplyr::distinct() %>%
  dplyr::group_by(wikidata_id) %>%
  dplyr::mutate(sideA = paste(belligerent1, collapse = ", ")) %>%
  dplyr::ungroup() %>%
  dplyr::select(wikidata_id, sideA) %>%
  dplyr::distinct()

df_sideB <- df_dbpedia %>%
  dplyr::select(wikidata_id, belligerent2) %>%
  dplyr::distinct() %>%
  dplyr::group_by(wikidata_id) %>%
  dplyr::mutate(sideB = paste(belligerent2, collapse = ", ")) %>%
  dplyr::ungroup() %>%
  dplyr::select(wikidata_id, sideB) %>%
  dplyr::distinct()

df <- dplyr::left_join(df, df_sideA)
df <- dplyr::left_join(df, df_sideB)

df_sf <- df %>% dplyr::filter(!is.na(lat)) %>%
  dplyr::filter(!is.na(lon))

## New column for means used
df_sf$means_aerialbombinglab[df_sf$means_aerialbombing == 1] <- "aerial bombing"
df_sf$means_airtoairlab[df_sf$means_airtoair == 1] <- "air to air"
df_sf$means_closeairsupportlab[df_sf$means_closeairsupport == 1] <- "close air support"
df_sf$means_cruisemissileslab[df_sf$means_cruisemissiles == 1] <- "cruise missiles"
df_sf$means_droneslab[df_sf$means_drones == 1] <- "drone strikes"
df_sf$means_groundtroopslab[df_sf$means_groundtroops == 1] <- "ground troops"
df_sf$means_paramilitarylab[df_sf$means_paramilitary == 1] <- "paramilitary"

df_sf$meanslab <- paste0(df_sf$means_aerialbombinglab, ", ", 
                         df_sf$means_airtoairlab, ", ", 
                         df_sf$means_closeairsupportlab, ", ", 
                         df_sf$means_cruisemissileslab, ", ", 
                         df_sf$means_droneslab, ", ", 
                         df_sf$means_groundtroopslab, ", ", 
                         df_sf$means_paramilitarylab)

# Delete NA and ", ," and padding
df_sf$meanslab <- gsub("NA, ", "", df_sf$meanslab)
df_sf$meanslab <- gsub(", NA", "", df_sf$meanslab)

# Create pop up label
## Pop up formatting
sep <- "<br>"
close_sep <- "</br>"
str_open <- "<strong>"
str_close <- "</strong>"

## Create label column
df_sf$label <- paste0(sep, str_open, "Operation: ", str_close, df_sf$wikidata_name, close_sep, 
                      sep, str_open, "Part of: ", str_close, df_sf$parent_name, close_sep, 
                      sep, str_open, "Dates: ", str_close, df_sf$start_date, " to ", df_sf$end_date, close_sep,
                      sep, str_open, "Side A: ", str_close, df_sf$sideA, close_sep,
                      sep, str_open, "Side B: ", str_close, df_sf$sideB, close_sep,
                      sep, str_open, "Means: ", str_close, df_sf$meanslab, close_sep)

# Coerce to sf object
df_sf <- sf::st_as_sf(df_sf, coords = c("lon", "lat"))

# Calculate center for view
center_lon <- mean(df$lon, na.rm = TRUE)
center_lat <- mean(df$lat, na.rm = TRUE)

# Plot with leaflet
leaflet::leaflet() %>%
  leaflet::addTiles() %>%
  leaflet::addCircleMarkers(data = df_sf,
                            popup = df_sf$label,
                            popupOptions = leaflet::popupOptions(closeButton = TRUE),
                            clusterOptions = leaflet::markerClusterOptions()) %>%
  leaflet::setView(center_lon, center_lat, zoom = 3)
```
